---
title: "Final"
author: "Liping He"
date: "12/5/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## R Markdown

```{r}
library(olsrr);library(DescTools); library(ResourceSelection);library(car)
```


```{r}
birthweight=read.csv("birthweight_final.csv", header=TRUE)
birthweight$Black = as.factor(birthweight$Black)
birthweight$Married = as.factor(birthweight$Married)
birthweight$Boy = as.factor(birthweight$Boy)
birthweight$MomSmoke = as.factor(birthweight$MomSmoke)
birthweight$Ed = as.factor(birthweight$Ed)
```

##Q1 
###Q1(1)
```{r}
#stepwise selection
#Black+Married+Boy+MomSmoke+Ed+MomAge+MomWtGain+Visit
full.model <- lm(Weight ~ .-Weight_Gr, data = birthweight)
model.stepwise<-ols_step_both_p(full.model, pent = 0.01, prem = 0.01, details = FALSE)
model.stepwise
#plot(model.stepwise)
```


```{r}
# forward selection
model.forward<-ols_step_forward_p(full.model, penter = 0.01, details = F) # penter: threshold p-value for enter
model.forward    # final model summary
#plot(model.forward)
```


```{r}
# backward selection
model.backward<-ols_step_backward_p(full.model, prem = 0.01, details = F) # prem: threshold p-value for removal
model.backward 
#plot(model.backward)
```

```{r}
#best subset selection (AIC, SBC, adjusted R-squre, C(p) etc.)
model.best.subset<-ols_step_best_subset(full.model) 
model.best.subset
```

###Q1(2)
```{r}
lm.birth.muti <- lm(Weight ~ MomWtGain+MomSmoke+Black, data = birthweight)
summary(lm.birth.muti)
```
```{r}
with(birthweight, plot(birthweight,Weight))
abline(lm.birth.muti, col="red")
par(mfrow=c(2,2))
plot(lm.birth.muti, which=c(1:4))
```


```{r}
cook.muti = cooks.distance(lm.birth.muti)
#round(cook.d2, 2)
plot(cook.muti,col="pink", pch=19, cex=1)
text(cooks.distance(lm.birth.muti),labels = rownames(birthweight))
```

```{r}
inf.muti=which(cooks.distance(lm.birth.muti)>0.115)
birthweight[inf.muti, ]
```

###Q1(2)
From the diagnostics plot, there are some outliers, and there is data with Cook’s distance larger than 0.115. 
But the majority data follow our model’s prediction line, means they are close to the actual values, on average.
All selected predictors are significant, so we do no need refit the model.

There are 3 observations in the final model,MomWtGain, MomSmoke and Black.

###Q1(3)
Multiple R-squared:  0.1271, there are 12.71% of the variation of Weight is explained by the final model.

###Q1(4)
An increase of 1 unit in MomWtGain predicts an increase of 12.006 units of Weight on average, when all other predictors are fixed. 
On average, there is 237.799 units lower for smoke mom on baby birth weight than non somke mom when MomGain and Black are same.
On average, there is 236.556 units lower for black mom on baby birth weight than white mom when MomGain and MomSmoke are same.


##Q2

###Q2(1)
```{r}
## model selection - setpwise selection
glm.null.Gr <- glm(Weight_Gr ~ 1, data = birthweight, family = "binomial")
glm.full.Gr <- glm(Weight_Gr ~ Black+Married+Boy+MomSmoke+Ed+MomAge+MomWtGain+Visit, data = birthweight, family = "binomial")

```

```{r}
# stepwise selection with AIC
step.models.AIC<-step(glm.null.Gr, scope = list(upper=glm.full.Gr),
                  direction="both",test="Chisq", trace = F) 
summary(step.models.AIC) 
```

```{r}
# stepwise selection with BIC
step.models.BIC<-step(glm.null.Gr, scope = list(upper=glm.full.Gr),
                  direction="both",test="Chisq", trace = F, k=log(nrow(birthweight))) 

summary(step.models.BIC)
```

###Q2(2)~(5)
```{r}
## goodness-of-fit check:  Hosmer Lemeshow test
hoslem.test(step.models.BIC$y, fitted(step.models.BIC), g=10) 
```

```{r}
#############################################
## residual plots
resid.d<-residuals(step.models.BIC, type = "deviance")
resid.p<-residuals(step.models.BIC, type = "pearson")
std.res.d<-residuals(step.models.AIC, type = "deviance")/sqrt(1 - hatvalues(step.models.BIC)) # standardized deviance residuals
std.res.p <-residuals(step.models.BIC, type = "pearson")/sqrt(1 - hatvalues(step.models.BIC)) # standardized pearson residuals

```


```{r}
dev.new(width = 1000, height = 1000, unit = "px")
par(mfrow=c(1,2))
plot(std.res.d[step.models.BIC$model$Weight_Gr==0], col = "red", 
     ylim = c(-3.5,3.5), ylab = "std. deviance residuals", xlab = "ID")
points(std.res.d[step.models.BIC$model$Weight_Gr==1], col = "blue")

plot(std.res.p[step.models.AIC$model$Weight_Gr==0], col = "red", 
     ylim = c(-3.5,3.5), ylab = "std. Pearson residuals", xlab = "ID")
points(std.res.p[step.models.BIC$model$Weight_Gr==1], col = "blue")

```

```{r}
## 2. Influence dianostics - cook's distance

dev.new(width = 1000, height = 1000, unit = "px")
plot(step.models.BIC, which = 4, id.n = 5)  # visual inspection

# which observation has cook's d larger than 0.1?
(inf.id=which(cooks.distance(step.models.BIC)>0.1))

inf.id
```


```{r}
round(exp(step.models.BIC$coefficients),3)
```


```{r}
## 1. predict the probability (p) 
fit.prob <- predict(step.models.BIC, type = "response") # estimated (fitted) probabilities of unhealthy status

## 2. classification with threshold
pred.class.1 <- ifelse(fit.prob > 0.5, 1, 0) # classification with mean threshold

# output in slide - not necessary but just for the display
sample.prop=mean(birthweight$Weight_Gr) # sample proportion
sample.prop
pred.class.2 <- ifelse(fit.prob > sample.prop, 1, 0) # classification with sample proportion threshold
combined.dat = cbind(birthweight[,c("MomWtGain","MomSmoke","MomAge")], fit.prob, pred.class.1,pred.class.2)
combined.dat[12:20,]
head(pred.class.2, 10)

## 3. calculate misclassification rate for each threshold
mean(birthweight$Weight_Gr != pred.class.1)  # misclassification rate from 0.5 threshold
mean(birthweight$Weight_Gr != pred.class.2)  # misclassification rate from sample proportion threshold
```
###Q2(2)
There are three variables from our final mode,MomWtGain,MomSmoke,MomAge, there are all significant based on significance level 0.05. So we don't need to re-fit the model.
Most observations standardized residuals fall between -2 and 2, so we say there is no serious issue.
There is no residuals cook's distance greater than 0.1.

###Q2(3)
The odds of being birthweight for a mom change by factors of:
exp(-0.036819)=0.964 with one unit increase in MomWtGain;
exp(-0.048266) = 0.953 with one unit increase in MomAge;
the odds of smoke mom 2.377 (=exp(0.865786)) times of a non-smoke mom 

###Q2(4)
A more weight gain, older and smoke woman has the high chance to deliver a low birthweight infant.

###Q2(5)
The sample proportion of low birthweight infant in the dataset is 0.4925.

###Q2(6)
The misclassification rate is 0.355

###Q2(7)
The goodness of fit test for the model has a p-value of 0.3252, is greater than significance level 0.05 
it indicates the model fit is reasonable.

##Q3
Compare the results from exercise 1-2, variables MomWtGain and MomSomke are significant to birthweight.    
Black is significant in linear regression while MomAge is significant in logical regression. 
Low birthweight is a risk factor that can lead infant mortality, I would suggest pregnant women control their weight growth, and do not smoke during pregnancy.
