---
title: "HW3"
author: "Liping He"
date: "10/20/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS); library(car); library(olsrr)
```

##Exercise 1:
#Q1
```{r}
setwd("/Users/lilyhe/Documents/UTSA fall 2021/STA6443/HW3")
heart = read.csv("heart.csv", header = TRUE)
str(heart)
head(heart,10)
```

```{r}
lm.heart <- lm(Cholesterol~Weight, data= heart)
summary(lm.heart)
```

```{r}
with(heart, plot(Weight,Cholesterol))
abline(lm.heart, col="red")
par(mfrow=c(2,2))
plot(lm.heart, which=c(1:4))
```

```{r}
cook.d = cooks.distance(lm.heart)
plot(cook.d,col="pink", pch=19, cex=1)
text(cooks.distance(lm.heart),labels = rownames(heart))
```

```{r}
inf.id=which(cooks.distance(lm.heart)>0.015)
heart[inf.id, ]
lm.heart2=lm(Cholesterol ~ Weight, data=heart[-inf.id, ])
```

```{r}
with(heart, plot(Weight, Cholesterol))
abline(lm.heart, col="red")
abline(lm.heart2,col="blue")
legend("bottomright",col=c("red","blue"),legend=c("w/ 23&210", "w/out 23&210"), cex=0.8, title.adj=0.15, lty=1)
```
```{r}
with(heart, plot(Weight,Cholesterol))
abline(lm.heart2, col="red")
par(mfrow=c(2,2))
plot(lm.heart2, which=c(1:4))
```



```{r}
summary(lm.heart)   # w/ 23&210 
summary(lm.heart2)  # w/out 23&210
```
#Q2

From the R output, Weight is statistically significat predictor since it's p-value smaller than significant level 0.05. From the F-statistic is slightly increased from 15.22 to 19.97 after removal of extreme values, it is greater than 0, so we say on average, Y is predicted to have an increase of 19.97 when x increases by 1 unit, so there is a relationship between our predictor variable (Weight) and response variable (Cholesterol), and the model is useful.We can formulate the model: y^ = 203.57605+ 0.12264*Weight

From below aspects, we say this is not a good model for the prediction of Cholesterol levels:
1) The R square we got is 0.006339 from the refit the model so we say roughly 0.6% of the variance found in the response variable Cholesterol can be explained by the Weight. Since it is very small,so, our model doesn't have good-of-fit of the data. 
2) The slop is 0.006339, and from the pot, the line almost horizontal, means the slop is not strong at all for Y by x.
3) From the polt, dots are mess and a lot scatters, they are far way from our model’s prediction line, means they are not close to the actual values, on average.
4）There are larger numbers of residuals with extreme values.
4) From the standardized residuals plot, some data larger than 2.


##Exercise 2

#a)
```{r}
lm.heart.muti <- lm(Cholesterol~ Diastolic + Systolic +Weight, data= heart)
summary(lm.heart.muti)
```

```{r}
cook.muti = cooks.distance(lm.heart.muti)
#round(cook.d2, 2)
plot(cook.muti,col="pink", pch=19, cex=1)
text(cooks.distance(lm.heart.muti),labels = rownames(heart))
```


```{r}
inf.muti=which(cooks.distance(lm.heart.muti)>0.015)
heart[inf.muti, ]
lm.heart.muti=lm(Cholesterol ~ Diastolic + Systolic +Weight, data=heart[-inf.muti, ])

```

```{r}
summary(lm.heart.muti)
```
```{r}
par(mfrow=c(2,2))
plot(lm.heart.muti, which=1:4)  # diagnostics plot 

```

##b)
We can use these coefficients to form the following estimated regression equation:
y^ = 156.32618+0.24922*Diastolic + 0.03671*Systolic+0.02146*Weight.Y is predicted to have an increase of 156.32618 when all other predictiors are fixed.
If we used a significant level of 0.05 to determine which predictors were significant in this regression model, we’d say that Diastolic and Systolic are statistically significant predictors while Weight is not.Since p-value is very small, we say at least one β>0, and the model is useful.
Multiple R-squared is 0.03606, so 3.6% of the variance found in the response variable Cholesterol can be explained by the our model. Since it is very small,so, our model doesn't have good-of-fit of the data.

##Exercise 3

```{r}

full.model <- lm(Weight ~., data = heart[-inf.muti, ])
model.stepwise<-ols_step_both_p(full.model, pent = 0.10, prem = 0.10, details = FALSE)
model.stepwise
plot(model.stepwise)

```

```{r}
lm.step=lm(Cholesterol ~ Diastolic + Systolic, data=heart[-inf.muti, ])
summary(lm.step)
```
##a),b)
The p-value of the model and individual variables are all smaller then significance level 0.05, so, the model is useful.We can formulate the model: y^ = 159.3317+ 0.2770*Diastolic+0.3022*Systolic
Compare the R output of stepwise and exercise 1 and 2, the stepwise model removed predictor variable Weight since it is insignificant, while in exercise it is significant.
The further the F-statistic is from 1 the better it is, we can see for the F-statistic stepwise(60.38) > exercise 2(40.81) > exercise 1(39.03).

##Exercise 4
```{r}
model.best.subset<-ols_step_best_subset(full.model)
model.best.subset
```

```{r}
lm.Rsquare=lm(Cholesterol ~ Diastolic + Systolic +Weight, data=heart[-inf.muti, ])
summary(lm.Rsquare)
```

```{r}
lm.AIC=lm(Cholesterol ~ Diastolic + Systolic, data=heart[-inf.muti, ])
summary(lm.AIC)
```


#a)If we select the best model based on adjusted_R square criteria, the larger the better, we select model 3. Selected predictors are:Weight,Diastolic and Systolic.
#b)If we select the best model based on AIC criteria, the smaller the better, we select model 2. Selected predictors are:Diastolic and Systolic.
#c) The final model in a) has 1 more predictor variable, which is Weight, but from the model, we can see it's p-value is 0.1994, which is greater than 0.05. Stepwise selection also got model 2 as final selection, so we think model 2 is better.
