---
title: "Homework 1"
author: "Liping He, llt949"
date: "07 September 2021"
output: html_document
---


```{r setup, include=FALSE}
options(pillar.sigfig = 3)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This homework uses `tech_co_cstat_dta.zip` which we used in weeks 2 and 3. The data set consists of the following variables:

```{r}
d1 <- haven::read_dta("tech_co_cstat_dta.zip")
names(d1)
```
Read `tech_co_cstat_dta.zip` into your R session using `read_dta()` function from `haven` package. Store the resulting object in `d1`.

```{r}
d1 <- haven::read_dta("tech_co_cstat_dta.zip")
```


Before you start working on this homework, study the variables in `tech_co_cstat_dta.zip` as well as the structure of the dataset by typing these commands in your console:

```{r echo = TRUE, eval=FALSE}
psych::describe(d1)
dplyr::glimpse(d1)
```

Take a peek at the data by typing:

```{r echo=TRUE, eval=FALSE}
head(d1)
```

Read the attributes of any variable from this data set using `attributes` function. For example, the attributes of `gvkey` can be printed using:

```{r echo=TRUE}
attributes(d1$gvkey)
```


Finally, before you begin, include only the rows with `sale > 0`.

```{r echo=TRUE}
d2 <- filter(d1, sale > 0)
```


<p style="border:3px; border-style:solid; border-color:#b50000; padding: 1em;">This homework consists of 8 questions. Q1 carries 1 point. Q2 through Q8 carry two points each. We use `d2` as the initial input. Your objective is to reproduce the output shown in the HTML file for Q1 through Q9. </p>


### Q1

Print a data frame with the medians of `cogs`, `emp`, and `xrd`. 


```{r}
select(d1, cogs, emp, xrd) %>% 
  summarise(cogs = median(d1$cogs, na.rm = TRUE), 
            emp = median(d1$emp,na.rm = TRUE),
            xrd = median(d1$xrd,na.rm = TRUE))
```
```{r}
#from teacher
d2 %>%
  summarize(across(c(cogs, emp, xrd), median, na.rm = TRUE))
```

### Q2

Print a data frame with the means of `sale`, `oibdp`, and `xrd` for Apple, Facebook, and Tesla. For this, you will need to follow these steps:

1. Filter only the observations pertaining to Apple, Facebook, and Tesla
2. Group by `conm`
3. Summarize across `sale`, `oibdp`, and `xrd` to get their means
4. Output it as data frame by using `as.data.frame()` function.

```{r}
d10<-
filter(select(d1,conm,sale, oibdp, xrd),grepl("APPLE",conm) | grepl("FACE",conm)| grepl("TESLA",conm)) %>% 
  group_by(conm) %>% 
  summarise(across(where(is.numeric), mean)) %>% 
  as.data.frame()
d10
```

```{r}
#from teacher
d2 %>% 
filter(conm %in% c("APPLE INC","FACEBOOK INC","TESLA INC")) %>% 
  group_by(conm) %>% 
  summarize(across(c(cogs, emp, xrd), median, na.rm = TRUE)) %>% 
  as.data.frame()
```

### Q3

Round all the numeric variables in the above data frame to 1 decimal place. Output as a data frame using `as.data.frame()` function.

For rounding, you will have to use `mutate`, `across`, and `where` functions from `dplyr` package. Check https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-colwise/ for more information.

```{r}
select(d10,conm, sale, oibdp, xrd) %>% 
  mutate_if(is.numeric, ~round(., 1)) %>% 
  as.data.frame() 

```
```{r}
#from teacher
d2 %>% 
filter(conm %in% c("APPLE INC","FACEBOOK INC","TESLA INC")) %>% 
  group_by(conm) %>% 
  summarize(across(c(cogs, emp, xrd), median, na.rm = TRUE)) %>% 
  mutate(across(where(is.numeric), round, 1)) %>% 
  as.data.frame()
```

### Q4

Many advertising values are missing. The missing code in R is `NA`. We can get the total number of missing values for advertising quite easily by running the following function:

```{r echo=TRUE}
sum(is.na(d2$xad))
```

In the finance literature, a common (but incorrect) practice is to assume that the missing advertising is 0. We will use this adjustment to `xad` and create a new variable `adv` and save it in a new object `d3`. 

The first six values of `d3` when `xad` is NA are as follows:

```{r}
d3 <- select(d2, conm, datadate, xad) %>% 
  mutate(adv=ifelse(is.na(d2$xad),0,xad))%>% 
  filter(is.na(d2$xad)) 
  
d3 %>% head(6)
```

```{r}
#from teacher
d3 <- d2 %>% 
  mutate(adv = ifelse(is.na(xad), 0, xad))

d3 %>% 
  filter(is.na(xad)) %>% 
  select(conm, datadate, xad, adv) %>% 
  head()
```

### Q5

Using `d3`, create the following variables and print first 8 rows for Netflix and the new columns along with `conm` and `datadate`:

1. Return on assets (`roa`) = `oibdp` / `at`
2. Free cash flow (`fcf`) = `oancf` / `che`
3. Strategic emphasis (`strat_emph`) = (`adv` - `xrd`) / `at`

```{r}
select(d2, conm, datadate,xad,oibdp,at,oancf,che,xrd)%>% 
  mutate(adv=ifelse(is.na(d2$xad),0,xad)) %>% 
  mutate(roa=ifelse(is.na(d2$oibdp),0,oibdp)/at) %>% 
  mutate(fcf=ifelse(is.na(d2$oancf),0,oancf)/che) %>% 
  mutate(strat_emph=(adv-xrd)/at)%>%
  filter(grepl("NETFLIX",conm)) %>% 
  select(conm,datadate,roa,fcf,strat_emph) %>% 
  mutate_if(is.numeric, ~round(., 4)) %>% 
  as.data.frame() %>% 
  head(8)
```

```{r}
#from teacher
d3 %>% 
  filter(conm == "NETFLIX INC") %>% 
  mutate(roa = oibdp / at,
         fcf = oancf / che,
         strat_emph = (adv -xrd) / at
         ) %>% 
  select(conm, datadate, roa, fcf, strat_emph) %>% 
  head(8)
```


### Q6

You want to know how many profitable years each of the sample company experienced. For this follow these steps:

1. Create an indicator variable (dummy variable) called `profit_ind` such that when `oibdp > 0` this variable is 1. Otherwise it is 0.
2. Group by company names
3. Summarize `profit_ind` by taking its sum. Also, get the total number of observations for each company.


```{r}
select(d2, conm, oibdp)%>% 
  mutate(profit_ind=ifelse(oibdp > 0,1,0)) %>% 
  group_by(conm) %>% 
  summarise(profit_years=sum(profit_ind),total_years=n()) 
```
```{r}
d3 %>% 
  mutate(profit_ind=ifelse(oibdp > 0,1,0)) %>% 
  group_by(conm) %>% 
  summarise(profit_years=sum(profit_ind),total_years=n()) 
```

### Q7

Find the average annual stock returns of all the companies. Follow these steps:

1. Arrange the data set by `conm` and `datadate`.
2. Group by `conm`
3. Calculate stock return `stk_ret` by taking the difference between `prcc_f` and its lag and then divide the difference by the lag of `prcc_f`.
4. Summarize to get the mean of the stock returns `stk_ret_mean`.
5. Display the average stock returns in percentage format. 


```{r}
select (d1,conm,datadate,prcc_f) %>% 
arrange(conm,datadate) %>%
  group_by(conm) %>%
  mutate(stk_ret =(prcc_f-lag(prcc_f))/ lag(prcc_f)) %>% 
  group_by(conm) %>% 
  summarise(stk_ret_mean =mean(stk_ret,na.rm=TRUE)) %>% 
  mutate(stk_ret_mean=paste0(round(stk_ret_mean*100,3),"%")) %>% 
  as.data.frame()
```

```{r}
d3 %>% 
  arrange(conm, datadate) %>% 
  group_by(conm) %>% 
  mutate(stk_ret = (prcc_f- lag(prcc_f)) / lag(prcc_f)) %>% 
  summarise(stk_ret_mean =mean(stk_ret,na.rm=TRUE)) %>%
  mutate(stk_ret_mean = scales::percent(stk_ret_mean)       
         )
```

### Q8

In many statistical and machine learning applications, we use scaled variables instead of the original variables. A scaled variable is typically created by subtracting the sample mean of the variable from the variable and dividing it by its standard deviation. There is a `scale()` function in base R which can directly do it. 

You want to create a scaled variable for `sale` but separately for each company. Therefore, you can't use the mean and standard deviation of `sale` for the entire sample. Instead, you have to calculate these statistics for each company separately and then create a scaled variable. Follow these steps:

1. Group by `conm`
2. Summarize `sale` to get the mean (`sale_mean`) and the standard deviation (`sale_sd`)
3. Assign this data frame to `d3_sum`
4. Join `d3` and `d3_sum` by `conm`
5. Create `sale_scaled` by subtracting `sale_mean` from `sale` and dividing this difference by `sale_sd`

Print the first 10 rows for Twitter with `conm`, `sale`, `sale_scaled`, `sale_mean`, and `sale_sd` using `as.data.frame()`

```{r}
d3_sum <- d2 %>% 
  select(conm,sale) %>% 
  group_by(conm) %>% 
  summarise(sale_mean = mean(sale),sale_sd = sd(sale))
```


```{r}
d3_sum %>% 
  right_join(select(d2,conm,sale) , by = "conm") %>% 
  mutate(sale_scale =(sale - sale_mean)/sale_sd) %>% 
  filter(grepl("TWITTER",conm)) %>% 
  select(conm,sale,sale_scale,sale_mean,sale_sd) %>% 
  as.data.frame()
```

```{r}
#from teacher
d3_sum <- d3 %>% 
  group_by(conm) %>% 
  summarise(sale_mean = mean(sale),sale_sd = sd(sale))
```


```{r}
d3 %>% 
  inner_join(d3_summ, by ="conm") %>% 
  mutate(sale_scale =(sale - sale_mean)/sale_sd) %>%
  filter(conm =="TWITTER INC",conm) %>% 
  select(conm,sale,sale_scale,sale_mean,sale_sd) %>% 
  as.data.frame()  
```


<hr style="border: 2px solid#5b9aa0;" />





