---
title: "Dow Jones Case Study"
author: "Michael Adebisi, Lily He, Sadiyah Hotakey, Justin Howard, Olamiposi Sunmola "
date: "3/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tseries)
```

```{r}
library(quantmod)
```

```{r}
dow <- read.csv("/Users/lilyhe/Documents/01 UTSA Spring 2022/02_DA6813/casestudy/casestudy03/dow_jones_index.data", header = TRUE)
```

```{r}
#Remove incomplete cases of data object. Find missing NA values
dow<-na.omit(dow)
anyNA(dow)
```

```{r}
#Compactly display internal structure of object
str(dow)
```

```{r}
#Factor variables & convert to numeric
dow$date = lubridate::mdy(dow$date)
dow$open = as.numeric(gsub("\\$", "", dow$open))
dow$high = as.numeric(gsub("\\$", "", dow$high))
dow$low = as.numeric(gsub("\\$", "", dow$low))
dow$close = as.numeric(gsub("\\$", "", dow$close))
dow$volume = as.numeric(dow$volume)
dow$next_weeks_open = as.numeric(gsub("\\$", "", dow$next_weeks_open))
dow$next_weeks_close = as.numeric(gsub("\\$", "", dow$next_weeks_close))
str(dow)
```

```{r}
library(dplyr)
```

```{r}
#Categorize data into groups by stock then induce lagged values for the variable.

#Open
dow=dow %>%
group_by(stock)%>%
  mutate(open.lag=dplyr::lag(open,n=1))

#High
dow=dow %>%
group_by(stock)%>%
  mutate(high.lag=dplyr::lag(high,n=1))

#Low
dow=dow %>%
group_by(stock)%>%
  mutate(low.lag=dplyr::lag(low,n=1))

#Close 
dow=dow %>%
group_by(stock)%>%
  mutate(close.lag=dplyr::lag(close,n=1))


#Percent_Change_Volume_Over_Last_Week 
dow=dow %>%
group_by(stock)%>%
  mutate(percent_change_volume_over_last_wk.lag=dplyr::lag(percent_change_volume_over_last_wk,n=1))

# Percent_Change_Next_Weeks_Price 
dow=dow %>%
group_by(stock)%>%
  mutate(percent_change_next_weeks_price.lag=dplyr::lag(percent_change_next_weeks_price,n=1))

#Percent_Return_Next_Dividend
dow=dow %>%
group_by(stock)%>%
  mutate(percent_return_next_dividend.lag=dplyr::lag(percent_return_next_dividend,n=1))

#Next_Weeks_Close_Lag
dow=dow %>%
group_by(stock)%>%
  mutate(next_weeks_close.lag=dplyr::lag(next_weeks_close,n=1))

#Next_Weeks_Open_Lag
dow= dow %>%
  group_by(stock) %>%
  mutate(next_weeks_open.lag = dplyr::lag(next_weeks_open, n =1))
```

```{r}
#Split data into train & test data sets.
newdata<-split(dow,dow$quarter)
train<-newdata[[1]]
test<-newdata[[2]]
train1<-split(train,train$stock)
test1<-split(test,test$stock)
```


```{r}
library(tree)
library(caret)
```

```{r}
#Decision Tree model
treefxn <- function(trainstock, teststock, formula){
  set.seed(123)
  treefit <- tree(formula, data = trainstock)
  tree.predict <- predict(treefit, newdata = teststock)
  tree.predict <- ifelse(tree.predict >= mean(tree.predict), 1,0)
  tree.obs <- ifelse(teststock$percent_change_next_weeks_price >= mean(teststock$percent_change_next_weeks_price ),1,0)
  table <- table(tree.obs, tree.predict); print(summary(treefit))
  accuracy <- sum(diag(table))/13
  print(accuracy)
}

formula = percent_change_next_weeks_price ~ open.lag + high.lag + low.lag + close.lag + next_weeks_open.lag + next_weeks_close.lag + percent_return_next_dividend.lag + volume

#Create for loop for each stock.
for (i in names(train1)){
x= train1[[i]]
y= test1[[i]]
treefxn(x,y, formula)
}
```

```{r}
# CAPM and Stock Risk and Reward Calculations
dow_jones_2 = read.csv("/Users/lilyhe/Documents/01 UTSA Spring 2022/02_DA6813/casestudy/casestudy03/dow_jones_index.data", header = TRUE)

# convert factors to numeric
dow_jones_2$open=as.numeric(gsub("\\$","",dow_jones_2$open))

dow_jones_2$high=as.numeric(gsub("\\$","",dow_jones_2$high))

dow_jones_2$low=as.numeric(gsub("\\$","",dow_jones_2$low))

dow_jones_2$close=as.numeric(gsub("\\$","",dow_jones_2$close))

dow_jones_2$next_weeks_open=as.numeric(gsub("\\$","",dow_jones_2$next_weeks_open))

dow_jones_2$next_weeks_close=as.numeric(gsub("\\$","",dow_jones_2$next_weeks_close))

#Creating a 30-element list, each element being a single stock
DowJData3 <- split(dow_jones_2, dow_jones_2$stock)

#Reading in reference data
SP500Dta <- read.csv("/Users/lilyhe/Documents/01 UTSA Spring 2022/02_DA6813/casestudy/casestudy03/^GSPC.csv")

#The reference data must match the actual Dow stock data in length
#Therefore, only the first 25 rows are extracted:
SP500Dta<-SP500Dta[1:25,]

```

```{r}
#Compute percent change (using Delt function) for the closing price - reference data
ReturnSP500 = na.omit(Delt(SP500Dta[,5]))

#Compute percent change (using Delt function) for the closing price for each of the stocks
ReturnAA = na.omit(Delt(DowJData3$AA[,7]))
ReturnAXP = na.omit(Delt(DowJData3$AXP[,7]))
ReturnBA = na.omit(Delt(DowJData3$BA[,7]))
ReturnBAC = na.omit(Delt(DowJData3$BAC[,7]))
ReturnCAT = na.omit(Delt(DowJData3$CAT[,7]))
ReturnCSCO = na.omit(Delt(DowJData3$CSCO[,7]))
ReturnCVX = na.omit(Delt(DowJData3$CVX[,7]))
ReturnDD = na.omit(Delt(DowJData3$DD[,7]))
ReturnDIS = na.omit(Delt(DowJData3$DIS[,7]))
ReturnGE = na.omit(Delt(DowJData3$GE[,7]))
ReturnHD = na.omit(Delt(DowJData3$HD[,7]))
ReturnHPQ = na.omit(Delt(DowJData3$HPQ[,7]))
ReturnIBM = na.omit(Delt(DowJData3$IBM[,7]))
ReturnINTC = na.omit(Delt(DowJData3$INTC[,7]))
ReturnJNJ = na.omit(Delt(DowJData3$JNJ[,7]))
ReturnJPM = na.omit(Delt(DowJData3$JPM[,7]))
ReturnKO = na.omit(Delt(DowJData3$KO[,7]))
ReturnKRFT = na.omit(Delt(DowJData3$KRFT[,7]))
ReturnMCD = na.omit(Delt(DowJData3$MCD[,7]))
ReturnMMM = na.omit(Delt(DowJData3$MMM[,7]))
ReturnMRK = na.omit(Delt(DowJData3$MRK[,7]))
ReturnMSFT = na.omit(Delt(DowJData3$MSFT[,7]))
ReturnPFE = na.omit(Delt(DowJData3$PFE[,7]))
ReturnPG = na.omit(Delt(DowJData3$PG[,7]))
ReturnT = na.omit(Delt(DowJData3$T[,7]))
ReturnTRV = na.omit(Delt(DowJData3$TRV[,7]))
ReturnUTX = na.omit(Delt(DowJData3$UTX[,7]))
ReturnWMT= na.omit(Delt(DowJData3$WMT[,7]))
ReturnVZ = na.omit(Delt(DowJData3$VZ[,7]))
ReturnXOM = na.omit(Delt(DowJData3$XOM[,7]))
```

```{r}
#combine the percent change for the closing price for both the reference and actual data
MyData = cbind(ReturnSP500,ReturnAA,ReturnAXP,
               ReturnBA, 
               ReturnBAC,
               ReturnCAT,
               ReturnCSCO,
               ReturnCVX,
               ReturnDD,
               ReturnDIS,
               ReturnGE,
               ReturnHD,
               ReturnHPQ,
               ReturnIBM,
               ReturnINTC,
               ReturnJNJ,
               ReturnJPM,
               ReturnKO,
               ReturnKRFT,
               ReturnMCD,
               ReturnMMM,
               ReturnMRK,
               ReturnMSFT,
               ReturnPFE,
               ReturnPG,
               ReturnT,
               ReturnTRV,
               ReturnUTX,
               ReturnWMT,
               ReturnVZ,
               ReturnXOM)
               
colnames(MyData) = c("SP500", "AA", "AXP", "BA", "BAC", "CAT", "CSCO", "CVX", "DD", "DIS", "GE", "HD", "HPQ", "IBM", "INTC", "JNJ", "JPM", "KO", "KRFT", "MCD", "MMM", "MRK", "MSFT", "PFE", "PG", "T", "TRV", "UTX", "WMT", "VZ", "XOM")
head(MyData)
```

```{r}
#Compute mean and standard deviation for each of the columns in the combined dataset
DataMean = apply(MyData, 2, mean)
DataSD = apply(MyData, 2, sd)

#Bind means and standard deviations into one object
cbind(DataMean, DataSD)
```

```{r}
#Predict percent changes for the closing price for each of the stocks with the reference (SP500) percent change in closing price

#1
lm.AA <- lm(AA~ SP500, data = as.data.frame(MyData))
#2
lm.AXP <- lm(AXP~ SP500, data = as.data.frame(MyData))
#3
lm.BA <- lm(BA~ SP500, data = as.data.frame(MyData))
#4
lm.BAC <- lm(BAC~ SP500, data = as.data.frame(MyData))
#5
lm.CAT <- lm(CAT~ SP500, data = as.data.frame(MyData))
#6
lm.CSCO <- lm(CSCO~ SP500, data = as.data.frame(MyData))
#7
lm.CVX <- lm(CVX~ SP500, data = as.data.frame(MyData))
#8
lm.DD <- lm(DD~ SP500, data = as.data.frame(MyData))
#9
lm.DIS <- lm(DIS~ SP500, data = as.data.frame(MyData))
#10
lm.GE <- lm(GE~ SP500, data = as.data.frame(MyData))
#11
lm.HD <- lm(HD~ SP500, data = as.data.frame(MyData))
#12
lm.HPQ <- lm(HPQ~ SP500, data = as.data.frame(MyData))
#13
lm.IBM <- lm(IBM~ SP500, data = as.data.frame(MyData))
#14
lm.INTC <- lm(INTC~ SP500, data = as.data.frame(MyData))
#15
lm.JNJ <- lm(JNJ~ SP500, data = as.data.frame(MyData))
#16
lm.JPM <- lm(JPM~ SP500, data = as.data.frame(MyData))
#17
lm.KO <- lm(KO~ SP500, data = as.data.frame(MyData))
#18
lm.KRFT <- lm(KRFT~ SP500, data = as.data.frame(MyData))
#19
lm.MCD <- lm(MCD~ SP500, data = as.data.frame(MyData))
#20
lm.MMM <- lm(MMM~ SP500, data = as.data.frame(MyData))
#21
lm.MRK <- lm(MRK~ SP500, data = as.data.frame(MyData))
#22
lm.MSFT <- lm(MSFT~ SP500, data = as.data.frame(MyData))
#23
lm.PFE <- lm(PFE~ SP500, data = as.data.frame(MyData))
#24
lm.PG <- lm(PG~ SP500, data = as.data.frame(MyData))
#25
lm.T <- lm(T~ SP500, data = as.data.frame(MyData))
#26
lm.TRV <- lm(TRV~ SP500, data = as.data.frame(MyData))
#27
lm.UTX <- lm(UTX~ SP500, data = as.data.frame(MyData))
#28
lm.WMT <- lm(WMT~ SP500, data = as.data.frame(MyData))
#29
lm.VZ <- lm(VZ~ SP500, data = as.data.frame(MyData))
#30
lm.XOM <- lm(XOM~ SP500, data = as.data.frame(MyData))
```

```{r}
#Retrieve beta coefficients for each of the stocks by accessing the regression summary 
#Higher beta - higher risk

#1
BetaDowAA <- summary(lm.AA)$coefficient[2,1]
paste("Beta of AA:" , BetaDowAA)

#2
BetaDowAXP <- summary(lm.AXP)$coefficient[2,1]
paste("Beta of AXP:" , BetaDowAXP)

```

SVM Section:
```{r}
library(e1071) ##library for SVM functions
```

Read in the data
```{r}
dow1 <- read.csv("/Users/lilyhe/Documents/01 UTSA Spring 2022/02_DA6813/casestudy/casestudy03/dow_jones_index.data")
View(dow1)
```


Add continuous Y variable, this is for simple coding for the predictor variable percent_change_next_weeks_price
```{r}
dow2 <- dow1 %>% 
mutate(y = dow1$percent_change_next_weeks_price)

```


Add buy variable, if percent_change_next_weeks_price > 0 = 1 else = 0
What we are doing is creating a binary predictor variable, 
1 indicates buy (percentage increase next weeks price)
0 indicates do not buy (percentage decrease in next weeks price)
```{r}
dow2 <- dow2 %>%
mutate(buy = ifelse(percent_change_next_weeks_price > 0,1,0))
str(dow2)
```


Change buy variable to a factor
```{r}
dow2$buy = as.factor(dow2$buy)
str(dow2$buy)
```


create week variable, instead of using date which is listed as a character variable
create a variable called 'week' to make it easy to code for the models
seq week 1 to week 750 (number of observatins/rows in the dataset)
```{r}
dow2 <- dow2 %>%
mutate(week = seq(from = 1, to = dim(dow2)[1], by = 1))
str(dow2)
```


Change chr variables to numeric
```{r}
dow2$open = as.numeric(gsub("\\$", "", dow2$open))
dow2$high = as.numeric(gsub("\\$", "", dow2$high))
dow2$low = as.numeric(gsub("\\$", "", dow2$low))
dow2$close = as.numeric(gsub("\\$", "", dow2$close))
dow2$volume = as.numeric(dow2$volume)
dow2$next_weeks_open = as.numeric(gsub("\\$", "", dow2$next_weeks_open))
dow2$next_weeks_close = as.numeric(gsub("\\$", "", dow2$next_weeks_close))
str(dow2)
```



Create training data set
```{r}
train_dow = dow2 %>%
  filter(quarter ==1)
str(train_dow)
```


Create test data set
```{r}
test_dow = dow2 %>%
  filter(quarter ==2)
str(test_dow)
```


Creating the SVM RBF formula
```{r}
rbf_form = buy ~ . -quarter - stock - date - y
```


This may take 10 mins or so to run
```{r}
#tuned = tune.svm(rbf_formy, data = train_dow, gamma = seq(0.01, .5, by = 0.01), cost = seq(.1, 5, by = .1))
#commented this line out to help run all lines of code
```


what is the optimal gamma and cost for the svm model?
gamma = 0.01
cost = 5
```{r}
#tuned$best.parameters
#commented this line out to help run all lines of code
```


Creating the rbf Svm with the best parameters 
The syntax for SVM from the package e1071 is as follows for the default kernel, which is radial basis 
function (RBF): svm(formula = , data = , gamma =, cost =)

```{r}
rbf_svm = svm(formula = rbf_form, data = train_dow, gamma = 0.01, cost = 5)
summary(rbf_svm)

```


Performance of the model, fit rbf_svm model on the test_dow dataset

```{r}
rbf_svm_predict = predict(rbf_svm, test_dow, type = "response")

```


Use the command "confusionMatrix" from the package caret to get accuracy of the
model prediction. For the confusionMatrix the variable has to be a factor (categorical variable)

```{r}
caret::confusionMatrix(as.factor(test_dow$buy), as.factor(rbf_svm_predict))
```

Creating the linear svm model
```{r}
svm_lin_form = buy ~.-quarter -stock -date -y
lin_svm = svm(formula = svm_lin_form, data = train_dow, kernel = "linear", cost = 5, scale = FALSE)
summary(svm_lin_form)
```


Performance of the model, fit lin_svm model on the BBBC_Test dataset
```{r}
lin_svm_predict = predict(lin_svm, test_dow, type = "response")
```

Use the command "confusionMatrix" from the package caret to get accuracy of the
model prediction. For the confusionMatrix the variable has to be a factor (categorical variable)

```{r}
caret::confusionMatrix(as.factor(test_dow$buy), as.factor(lin_svm_predict))
```
